---
title: "Data manipulatiom"
output: github_document
---
```{r}
library(tidyverse)
```

## Load in the FAS litters data

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

choose some columns and not others.


Here, select which columns you want displayed.
```{r}
select(litters_df, group, litter_number)
```


Below, select columns from a starting point to an ending point (range) using a colon. 

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```





To get rid of a column, use a minus sign:
```{r}
select(litters_df, -litter_number)
```

Renaming columns.....
```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```
If you do not want to simultaneously select and rename,
```{r}
rename(litters_df, Group = group, Litter_NUMBER = litter_number)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```
these helper functions give us a quick way to select things - this selects everything that starts with 'gd'

```{r}
select(litters_df, litter_number, everything())
```
moves litter number to the beginning, but keeps everything else
another way to do this:
```{r}
relocate(litters_df, litter_number)
```

## 'filter'
keep only certain ROWS. ('select' operates on columns)

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```
! indicates that it is not equal to. 
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```
Tells us where it has con7 or mod8. Keep track of everything you want to include in your dataset!!


## 'mutate'

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```
Create a new variable, modify an existing variable. Everything in group is lower case, new variables get tacked on to the end. 

## 'arrange'
Function: put things in order

```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```
first in order for pups born alive, and within that, gestational day 0 weight.


## '%>%'
PIPING!!!!!!!!!

```{r}
litters_data_raw = read.csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight )
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

USE THE PIPE OPERATOR INSTEAD OF ALL OF THIS. Below yields the same results, but it's much cleaner. 

```{r}
litters_df = 
  read.csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)

```



