---
title: "Data manipulatiom"
output: github_document
---
```{r}
library(tidyverse)
```

## Load in the FAS litters data

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

choose some columns and not others.


Here, select which columns you want displayed.
```{r}
select(litters_df, group, litter_number)
```


Below, select columns from a starting point to an ending point (range) using a colon. 

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```





To get rid of a column, use a minus sign:
```{r}
select(litters_df, -litter_number)
```

Renaming columns.....
```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```
If you do not want to simultaneously select and rename,
```{r}
rename(litters_df, Group = group, Litter_NUMBER = litter_number)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```
these helper functions give us a quick way to select things - this selects everything that starts with 'gd'

```{r}
select(litters_df, litter_number, everything())
```
moves litter number to the beginning, but keeps everything else
another way to do this:
```{r}
relocate(litters_df, litter_number)
```

